### 2022-09-15 gloss.tex

\newacronym{chla}{Chl-\emph{a}}{Chlorophyll-\emph{a}}
\newacronym{sst}{SST}{Sea Surface Temperature}

\newacronym{hi}{HI}{Heterogeneity Index}

\newacronym{avhrr}{AVHRR}{Advanced Very-High-Resolution Radiometer}
\newacronym{modis}{MODIS}{Moderate Resolution Imaging Spectroradiometer}
\newacronym{viirs}{VIIRS}{Visible Infrared Imaging Radiometer Suite}