### 2023-01-10 gloss.tex
Up

\newacronym{chla}{Chl-\emph{a}}{Chlorophyll-\emph{a}}
\newacronym{SST}{SST}{Sea Surface Temperature}

\newacronym{HI}{HI}{Heterogeneity Index}

\newacronym{AVHRR}{AVHRR}{Advanced Very-High-Resolution Radiometer}
\newacronym{VIIRS}{VIIRS}{Visible Infrared Imaging Radiometer Suite}
\newacronym{MODIS}{MODIS}{Moderate Resolution Imaging Spectroradiometer}
\newacronym{ATSR}{ATSR}{(Advanced) Along Track-Scanning Radiometer}

\newacronym{NOAA}{NOAA}{National Oceanic and Atmospheric Administration}
\newacronym{ERS}{ERS}{European Remote Sensing}
\newacronym{MetOp}{MetOp}{Meteorological Operational satellite}

\newacronym{GHRSST}{GHRSST}{Group for High Resolution Sea Surface Temperature}
\newacronym{MUR}{MUR}{Multiscale Ultrahigh Resolution}

\newacronym{ESA}{ESA}{European Space Agency}
\newacronym{CCI}{CCI}{Climate Change Initiative}
\newacronym{C3S}{C3S}{Copernicus Climate Change Service}
\newacronym{OSTIA}{OSTIA}{Operational Sea Surface Temperature and Sea Ice Analysis}

\newacronym{CMEMS}{CMEMS}{Copernicus Marine Environment Monitoring Service}
\newacronym{CF}{CF}{Climate Forecast (convention for NetCDF)}