### 2023-01-21 tex/links.tex
Move glossaries

\newlink{gitlab}{gitlab}{gitlab.in2p3.fr/clement.haeck/submeso-color}
\newlink[doi=10.5281/zenodo.7470199]{zenodo}{Zenodo}{doi.org/10.5281/zenodo.7470199}

\newlink{numpy}{NumPy}{numpy.org}
\newlink{scipy}{SciPy}{scipy.org}
\newlink{pandas}{pandas}{pandas.pydata.org}

\newlink{xarray}{Xarray}{xarray.dev}
\newlink{dask}{Dask}{www.dask.org}

\newlink{xhistogram}{xhistogram}{github.com/xgcm/xhistogram}
\newlink{dask-histogram}{dask-histogram}{github.com/dask-contrib/dask-histogram}
\newlink{boost}{Boost Histogram}{github.com/scikit-hep/boost-histogram}

\newlink{matplotlib}{Matplotlib}{matplotlib.org}
\newlink{cartopy}{Cartopy}{scitools.org.uk/cartopy/docs/latest/}
\newlink{cmocean}{cmocean}{matplotlib.org/cmocean/}

\newlink{ipython}{IPython}{github.com/ipython/ipython}
\newlink{mamba}{Mamba}{github.com/mamba-org/mamba}
\newlink{conda-forge}{Conda-Forge}{conda-forge.org}
\newlink{doom}{Doom Emacs}{github.com/doomemacs/doomemacs}

\newlink{filefinder}{FileFinder}{github.com/Descanonge/filefinder}
\newlink{xarray-histogram}{Xarray-Histogram}{github.com/Descanonge/xarray-histogram}
\newlink{tol-colors}{Tol-Colors}{github.com/Descanonge/tol_colors}
\newlink{dateloop}{Dateloop}{github.com/Descanonge/dateloop}

\newlink{cf-xarray}{CF-Xarray}{github.com/xarray-contrib/cf-xarray}