### 2023-02-01 glossaries/links.tex
Cleaner glossaries and software in biblatex

\newlink{gitlab}{gitlab}{https://gitlab.in2p3.fr/clement.haeck/submeso-color}
\newdoi{zenodo}{Zenodo}{10.5281/zenodo.7470199}

\newlink{rv_histogram}{\texttt{scipy.stats.rv\_histogram}}{https://docs.scipy.org/doc/scipy/reference/generated/scipy.stats.rv_histogram.html}

\newlink{paultol}{Paul Tol}{https://personal.sron.nl/~pault}