### 2023-02-02 glossaries/links.tex

\newlink{gitlab}{gitlab}{https://gitlab.in2p3.fr/clement.haeck/submeso-color}
\newdoi{zenodo}{Zenodo}{10.5281/zenodo.7470199}

\newlink{rv_histogram}{\texttt{scipy.stats.rv\_histogram}}{https://docs.scipy.org/doc/scipy/reference/generated/scipy.stats.rv_histogram.html}

\newlink{paultol}{Paul~Tol}{https://personal.sron.nl/~pault}

\newlink{epsg-chl}{\texttt{EPSG:32662}}{https://epsg.io/32662}
\newlink{epsg-sst}{\texttt{EPSG:4326}}{https://epsg.io/4326}