### 2023-02-21 glossaries/links.tex
Fix citations in article

\newlink{gitlab}{gitlab}{https://gitlab.in2p3.fr/clement.haeck/submeso-color}
\newdoi{zenodo}{Zenodo}{10.5281/zenodo.7470199}

\newlink{biogeosciences}{Biogeosciences}{https://www.biogeosciences.net/}
\newlink{preprint-bgs}{Pre-print}{https://egusphere.copernicus.org/preprints/2023/egusphere-2022-1489/}
\newdoi{preprint-bgs-doi}{Pre-print}{10.5194/egusphere-2022-1489}

\newlink{rv_histogram}{\texttt{scipy.stats.rv\_histogram}}{https://docs.scipy.org/doc/scipy/reference/generated/scipy.stats.rv_histogram.html}

\newlink{paultol}{Paul~Tol}{https://personal.sron.nl/~pault}

\newlink{epsg-chl}{\texttt{EPSG:32662}}{https://epsg.io/32662}
\newlink{epsg-sst}{\texttt{EPSG:4326}}{https://epsg.io/4326}

\newlink{nco-packed}{NCO User Guide -- Packed data}{https://nco.sourceforge.net/nco.html\#Packed-data} %chktex 8