### 2023-03-10 glossaries/acronyms.tex
Up réanalyses

\newacronym{avhrr}{AVHRR}{Advanced Very"-High"-Resolution Radiometer}
\newacronym{viirs}{VIIRS}{Visible Infrared Imaging Radiometer Suite}
\newacronym{modis}{MODIS}{Moderate Resolution Imaging Spectroradiometer}
\newacronym[sort=AATSR]{atsr}{(A)ATSR}{(Advanced) Along Track"-Scanning Radiometer}
\newacronym{slstr}{SLSTR}{Sea and Land Surface Temperature Radiometer}
\newacronym{seawifs}{SeaWiFS}{Sea-viewing Wild Field"-of"-view Sensor}
\newacronym{meris}{MERIS}{Medium Resolution Imaging Spectrometer}
\newacronym{olci}{OLCI}{Ocean and Land Colour Instrument}
\newacronym[description={Joint Polar Satellite System, JPSS-1 = NOAA-20}]
{jpss}{JPSS}{Joint Polar Satellite System}

\newacronym{noaa}{NOAA}{National Oceanic and Atmospheric Administration}
\newacronym{ers}{ERS}{European Remote Sensing}
\newacronym{metop}{MetOp}{Meteorological Operational satellite}
\newacronym{snpp}{Suomi~NPP}{Suomi National Polar"-orbiting Partnership}

\newacronym{ghrsst}{GHRSST}{Group for High Resolution Sea Surface Temperature}
\newacronym{mur}{MUR}{Multiscale Ultrahigh Resolution}
\newacronym{cora}{CORA}{Coriolis Ocean dataset for ReAnalysis}

\newacronym{esa}{ESA}{European Space Agency}
\newacronym{cci}{CCI}{Climate Change Initiative}
\newacronym{c3s}{C3S}{Copernicus Climate Change Service}
\newacronym{ostia}{OSTIA}{Operational Sea Surface Temperature and Sea Ice Analysis}
\newacronym{ocssw}{OCSSW}{OceanColor Science Software}
\newacronym{ecmwf}{ECMWF}{European Centre for Medium"-Range Weather Forecast}

\newacronym{cmr}{CMR}{Common Metadata Repository}
\newacronym{obpg}{OBPG}{Ocean Biology Processing Group}
\newacronym{cmems}{CMEMS}{Copernicus Marine Environment Monitoring Service}
\newacronym{cf}{CF}{Climate Forecast (convention NetCDF)}
\newacronym{daac}{DAAC}{Distributed Active Archive Center}

\newacronym{pft}{PFT}{Phytoplanktonic Function Type}