### 2023-04-18 glossaries/links.tex
Add conferences

\newlink{gitlab}{gitlab}{https://gitlab.in2p3.fr/clement.haeck/submeso-color}
\newdoi{zenodo}{Zenodo}{10.5281/zenodo.7470199}

\newlink{biogeosciences}{Biogeosciences}{https://www.biogeosciences.net/}
\newlink{preprint-bgs}{Pre-print}{https://egusphere.copernicus.org/preprints/2023/egusphere-2022-1489/}
\newdoi{preprint-bgs-doi}{Pre-print}{10.5194/egusphere-2022-1489}

\newlink{rv_histogram}{\texttt{scipy.stats.rv\_histogram}}{https://docs.scipy.org/doc/scipy/reference/generated/scipy.stats.rv_histogram.html}

\newlink{paultol}{Paul~Tol}{https://personal.sron.nl/~pault}

\newlink{epsg-chl}{EPSG:32662}{https://epsg.io/32662}
\newlink{epsg-sst}{EPSG:4326}{https://epsg.io/4326}

\newlink{nco-packed}{NCO User Guide -- Packed data}{https://nco.sourceforge.net/nco.html\#Packed-data} %chktex 8

\newlink{cf-packed}{Conventions CF:\ Packed Data}
{http://cfconventions.org/Data/cf-conventions/cf-conventions-1.10/cf-conventions.html#packed-data}
\newlink{cf-flags}{Conventions CF:\ Flags}
{http://cfconventions.org/Data/cf-conventions/cf-conventions-1.10/cf-conventions.html#flags}

\newlink{mercator}{Mercator}{https://www.mercator-ocean.eu/}

\newlink{ve-illustration}{Visible Earth}{https://visibleearth.nasa.gov/images/146412/hints-of-spring-in-the-atlantic}
\newlink{eo-illustration}{Earth Observatory}{https://earthobservatory.nasa.gov/images/146412/hints-of-spring-in-the-atlantic}

\newlink{eo-sms}{Earth Observatory}{https://earthobservatory.nasa.gov/images/86449/blooming-baltic-sea}

\newlink{conf-filachange}{Filachange Workshop}{https://www.swot-adac.org/news-and-events/filachange-workshop}
\newlink{conf-egu22}{EGU 2022}{https://meetingorganizer.copernicus.org/EGU22/EGU22-8668.html}
\newlink{conf-carbon}{Ocean Carbon from Space Workshop}{https://oceancarbonfromspace2022.esa.int/}
\newlink{conf-egu21}{EGU 2021}{https://meetingorganizer.copernicus.org/EGU21/EGU21-4178.html}
\newlink{conf-lapcod}{LAPCOD 2019}{https://lapcod.rsmas.miami.edu/2019-Venice/index.php?page=home}