### 2023-04-23 tex/conclusion.tex
Up perspectives

% chktex-file 13

\chapter{Conclusion}
\label{chp:conclusion}
% plan à préciser